Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-09-11 main_new.tex
updates

Data are in our setting gathered by an autonomous vehicle that can measure
temperature and salinity at a subset of locations on the defined grid. This can be done in a static manner, where the survey design is pre-planned, or in a sequential way with several rounds of data gathering. For the latter, a round of data is gathered first, and then assimilated to get an updated model. Then the updated model is used to design and select the next round of measurements, and so on. 
We here describe the model for one round of data and the Gaussian formula for one round of updating. In Section
\ref{sec:heuristics} we discuss sequential data gathering in the context of adaptive design of AUV measurements.

Denote the data at $n_y$ locations by
$\by=(y_{a,1},y_{b,1},\ldots,y_{a,n_y},y_{b,n_y})^t$, representing the
temperature and salinity measurements gathered at one round grid nodes. The conditional model for the data, given the true temperature and salinity, is described by
\begin{equation}\label{likelihood}
\by | \bxi \sim N( \bA \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
\end{equation}
where $\bA$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
measurement indices on the grid, otherwise '$0$'. The size
$2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries $r^2_a$ and
$r^2_b$ which are measurement error variances of temperature and
salinity observations.

The GP mean and covariance in Eq. (\ref{prior}) are moderated when more information gets available. 
The updated distribution for temperature and salinity variables on the
spatial grid, given survey data $\by$, is Gaussian with mean and
covariance 
\begin{eqnarray}\label{gp_upd}
\bm &=& \bm(\by) = \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
\bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA \bSigma.\nonumber
\end{eqnarray}

The updated bivariate distribution at a grid location $\bx \in \mathcal{M}_g$ is then 
\begin{equation}\label{gp_hat}
\begin{bmatrix}
\xi_a(\bx) \\
\xi_b(\bx)
\end{bmatrix}
 |\by
 \sim N \left( 
\begin{bmatrix} m_{a}(\bx)\\
m_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
s_{a}^2(\bx) & s_{a,b}(\bx)  \\
s_{a,b}(\bx)  & s_{b}^2(\bx)  
\end{bmatrix}
\right),
\end{equation}
with entries extracted from the conditional mean and covariance
expressions in Eq. (\ref{gp_upd}).

>>>>

Data is gathered by an AUV that can measure temperature and salinity
at a subset of locations on the defined grid. This can be done in a
static manner, where the survey design is pre-planned, or in a
sequential way with several rounds of data gathering. For the latter,
a round of data is gathered first, and then assimilated to get an
updated model. Then the updated model is used to design and select the
next round of measurements, and so on. We describe the model for one
round of data and the Gaussian formula for a single update. In Section
\ref{sec:heuristics} we discuss sequential data gathering in the
context of adaptive design of AUV measurements.

We denote the data at $n_y$ locations by
$\by=(y_{a,1},y_{b,1},\ldots,y_{a,n_y},y_{b,n_y})^t$, representing the
temperature and salinity measurements gathered at the first round of
nodes in the grid. The conditional model for the data, given the true temperature
and salinity, is described by: 
\begin{equation}\label{likelihood}
\by | \bxi \sim N( \bA \bxi, \bR), %\pi(\by | \bxi)= \frac{1}{(2\pi)^{m}|\bR|^{\frac{1}{2}}}e^{-\frac{1}{2}(\by -\bA\bxi)^T\bR^{-1}(\by-\bA\bxi)},
\end{equation}
where $\bA$ is an $2n_y \times 2n$ matrix holding '$1$'s on the
measurement indices on the grid and '$0$' otherwise. The size
$2n_y \times 2n_y$ covariance matrix $\bR$ contains diagonal entries
$r^2_a$ and $r^2_b$ which are measurement error variances of
temperature and salinity observations.

The GP mean and covariance in Eq. (\ref{prior}) are moderated \kc{What
  does 'moderated' mean?} when more information is available. The
updated distribution for temperature and salinity variables on the
spatial grid, given survey data $\by$, is Gaussian with mean and
covariance: 
\begin{eqnarray}\label{gp_upd}
  \bm &=& \bm(\by) = \bmu+\bSigma \bA^t (\bA \Sigma \bA^t+\bR)^{-1}(\by-\bA \bmu),  \\
  \bS &=& \bSigma - \bSigma \bA^t (\bA \bSigma \bA^t+\bR)^{-1} \bA
          \bSigma.\nonumber
\end{eqnarray}

The updated bivariate distribution at a grid location $\bx \in
\mathcal{M}_g$ is then: 
\begin{equation}\label{gp_hat}
\begin{bmatrix}
\xi_a(\bx) \\
\xi_b(\bx)
\end{bmatrix}
 |\by
 \sim N \left( 
\begin{bmatrix} m_{a}(\bx)\\
m_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
s_{a}^2(\bx) & s_{a,b}(\bx)  \\
s_{a,b}(\bx)  & s_{b}^2(\bx)  
\end{bmatrix}
\right),
\end{equation}
with entries extracted from the conditional mean and covariance
expressions in Eq. (\ref{gp_upd}).
